Wireless Automobile Detection, License Plate Processing, and Data Availability Network Proposal

### 2014-03-17 proposal/proposal.tex
Final edits Sunday night

\subsection{Related Work}
The ACMx group at CSM has designed a system named SmartLots to track the number of vehicles in a parking lot. Their system is a wireless sensor network consisting of sensor nodes and a central base station. The sensor nodes are based on the Arduino Fio microcontroller platform equipped with triaxial magnetometers and XBee Pro radios. The base station is one small, commercially available Raspberry Pi Linux computer. The information collected by the sensor nodes is transmitted to the base station and forwarded to a server running a small web application. The system is described in \cite{stillwell2013} and \cite{parkingWiki}. \cite{stillwell2013} describes an implementation of a system to detect automobiles entering and exiting a parking lot using the Arduino Fio and the Raspberry Pi. This system uses the IEEE 802.15.4 communications standard to communicate automobile detection data from the sensor nodes to the base station. \cite{parkingWiki} describes the ongoing status of the ACMx project. At the time of writing this proposal, the ACMx team has created a website that will display the status of the CTLM upper and lower lots once the monitoring system has been deployed. \cite{parkingWiki} is updated periodically with new information about project progress. The project described in this proposal will augment and extend the ACMx project while collaborating closely with Stillwell (ACMx president and author of \cite{stillwell2013}), with the goal of a real world deployment by May 2014.

@@

The system will rely on three main components:
\begin{enumerate}
\item Detection Subsystem
\item Central Base Station
\item Server
\end{enumerate}
The detection subsystem will be made up of a small number of sensor nodes, placed strategically at the parking lot entrances. Each sensor node will contain an Arduino Fio equipped with a triaxial magnetometer that will be used to detect when a car passes the sensor. The work to determine the hardware configuration and software for the Arduino Fios has already been done by the ACMx group. Our contribution will be to also equip each node with a Raspberry Pi mini-computer and camera, which will take pictures of the license plates of cars as they pass by. When a car passes a sensor node, the magnetometer will detect the car, and the Arduino Fio will send an interrupt to the Raspberry Pi via the XBee radio, waking it up so it can take a picture of the license plate.

The central base station will be a Raspberry Pi placed in an upper window of the CTLM building, which will receive all the data from the sensor nodes via transmission from the XBee Pro radios attached to the sensor nodes. This base station Raspberry Pi will also have a camera attached that will take periodic images of the lots from above. The base station will forward the data to the ACMx server.

@@

\subsection{Automobile Detection}
The automobile detection subsystem provides a means for detecting ingress and egress of vehicles from a parking lot. This subsystem is to be placed on the side of the road next to each parking lot entrance and exit. Automobiles will be detected using a magnetometer which perceives the induced change in the local magnetic field as the metallic structure of the vehicle passes by, as described in \cite{stillwell2013}. The automobile detection subsystem will be based around the commercially available Arduino Fio 16-bit  platform which utilizes the ubiquitous Atmel ATMEGA328p microcontroller. When an automobile is detected by the Arduino Fio, a packet is sent through the onboard XBee Pro radio to the central base station (described in section 2.4). The packet will  contain the direction of the passing vehicle. At the time of detection,  a hardware interrupt is also sent to the license plate recognition system to awaken it from sleep.

@@

The hardware to be used for automobile detection will be contained in the same enclosure as the license plate image acquisition hardware for simplicity. Both components will share a unified power supply based on a high-capacity lithium polymer battery which is charged by a relatively small photovoltaic panel as seen in Figure~\ref{fig:autodetect}. A 5V boost converter will be required to interface with the Acquisition Raspberry Pi since the lithium polymer batteries only provide 3.7V.

The use of a magnetometer ensures that automobiles and other roadway vehicles such as motorcycles are detected, while pedestrians and particulate deposits, such as snow, are ignored. As in \cite{stillwell2013}, we shall use the commercially available HMC5883 triaxial magnetometer as it provides an adequate balance of sensitivity and low power consumption.

@@

The central base station, as described in section 2.4, will be placed in an upper window of the CTLM building, allowing it to take a clear overhead picture of at least one of the two parking lots. For the first deployment we may only have one overhead camera to start off with, although ideally we would have two - one for each lot. Several potential locations exist which allow for a clear line-of-sight to all automobile detection subsystems, one of which is detailed in Figure \ref{fig:deployment} The location of the base station must put it in range of the XBee Pros mounted on each sensor node, to allow data to be transferred from the detection subsystem to the base station. The locations being considered must also provide internet access to the base station Raspberry Pi through ethernet ports.

>>>>

\subsection{Related Work}
The ACMx group at CSM has designed a system named SmartLots to track the number of vehicles in a parking lot. Their system is a wireless sensor network consisting of sensor nodes and a central base station. The sensor nodes are based on the Arduino Fio microcontroller platform equipped with triaxial magnetometers and XBee Pro radios. The base station will be a Raspberry Pi connected to the internet via ethernet. The information collected by the sensor nodes is transmitted to the base station and forwarded to a server running a small web application. The system is described in \cite{stillwell2013} and \cite{parkingWiki}. \cite{stillwell2013} describes an implementation of a system to detect automobiles entering and exiting a parking lot using the Arduino Fio and the Raspberry Pi. This system uses the IEEE 802.15.4 communications standard to communicate automobile detection data from the sensor nodes to the base station. \cite{parkingWiki} describes the ongoing status of the ACMx project. At the time of writing this proposal, the ACMx team has created a website that will display the status of the CTLM upper and lower lots once the monitoring system has been deployed. \cite{parkingWiki} is updated periodically with new information about project progress. The project described in this proposal will augment and extend the ACMx project while collaborating closely with Stillwell (ACMx president and author of \cite{stillwell2013}), with the goal of a real world deployment by May 2014.

@@

The system will rely on three main components:
\begin{enumerate}
\item Detection Subsystem
\item Central Base Station
\item Server
\end{enumerate}
The detection subsystem will be made up of a small number of sensor nodes, placed strategically at parking lot entrances. Each sensor node will contain an Arduino Fio equipped with a triaxial magnetometer that will be used to detect when a car passes the sensor. The work to determine the hardware configuration and software for the Arduino Fios has already been done by the ACMx group. Our contribution will be to also equip each node with a Raspberry Pi mini-computer and camera, which will take pictures of the license plates of cars as they pass by. When a car passes a sensor node, the magnetometer will detect the car, and the Arduino Fio will send an interrupt to the Raspberry Pi via the XBee radio, waking it up so it can take a picture of the license plate.

The central base station will be a Raspberry Pi that will receive all the data from the sensor nodes via transmission from the XBee Pro radios attached to the sensor nodes. The base station Raspberry Pi will be centrally located in a nearby building with internet connectivity, and will have a camera attached that will take periodic images of the lots from above. The base station will use its network connection to forward the data to the ACMx server.

@@

\subsection{Automobile Detection}
The automobile detection subsystem will provide a means for detecting ingress and egress of vehicles from a parking lot. This subsystem is to be placed on the side of the road next to each parking lot entrance and exit. Automobiles will be detected using a magnetometer which perceives the induced change in the local magnetic field as the metallic structure of the vehicle passes by, as described in \cite{stillwell2013}. The automobile detection subsystem will be based around the commercially available Arduino Fio 16-bit  platform which utilizes the ubiquitous Atmel ATMEGA328p microcontroller. When an automobile is detected by the Arduino Fio, a packet will be sent through the onboard XBee Pro radio to the central base station (described in section 2.4). The packet will  contain the direction of the passing vehicle. At the time of detection, a hardware interrupt will also be sent to the license plate recognition system to awaken it from sleep mode.

@@

The hardware to be used for automobile detection will be contained in the same enclosure as the license plate image acquisition hardware for simplicity. Both components will share a unified power supply based on a high-capacity lithium polymer battery, which will be charged by a relatively small photovoltaic panel as seen in Figure~\ref{fig:autodetect}. A 5V boost converter will be required to interface with the Acquisition Raspberry Pi since the lithium polymer batteries only provide 3.7V.

The use of a magnetometer ensures that automobiles and other roadway vehicles such as motorcycles are detected, while pedestrians and particulate deposits, such as snow, are ignored. As mentioned in \cite{stillwell2013}, the automobile detection subsystem shall use the commercially available HMC5883 triaxial magnetometer as it provides an adequate balance of sensitivity and low power consumption.

@@

The central base station, as described in section 2.4, will be placed in an upper window of the CTLM building, allowing it to take a clear overhead picture of at least one of the two parking lots. For the first deployment we may only have one overhead camera to start off with, although ideally we would have two - one for each lot. Several potential locations exist which allow for a clear line-of-sight to all automobile detection subsystems, one of which is detailed in Figure \ref{fig:deployment}. The location of the base station must put it in range of the XBee Pros mounted on each sensor node, to allow data to be transferred from the detection subsystem to the base station. The locations being considered must also provide internet access to the base station Raspberry Pi through ethernet ports.
